---
title: "ARGs in WORLE poultry manure, 2017 wafergen results using Bob's primers"
author: "Jared"
date: "9/23/2019"
output:
  html_document:
    df_print: paged
    fig_width: 8.5
    fig_height: 11
    fig_caption: true
    toc: true
    toc_float: true
    number_sections: false
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=F, message=F}
library(tidyverse)
library(ggplot2)
library(reshape2)
library(ggpubr)
library(data.table)
```
    
This page details processing results from wafergen runs. In this case, the run was on the poultry manure used for the 
WORLE 2017 rainfall simulation with probes from "Bob's Primers".  
   
#### Read in results from wafergen, this is a tab seperated text file  
We can remove rows with Ct values that have NA as these were not amplified, test to see if any NAs in Ct column:
```{r, message=FALSE}
STRIPs <- read.delim("./data/wafergen_results/STRIPs.txt", row.names=NULL, comment.char="#")
# For this data, any rows with Ct == NA are not valuable, they indicated that a target sequence was not amplified in a sample so we can leave them out of this data

STRIPs <- STRIPs %>%
  filter(Ct != "NA")
anyNA(STRIPs$Ct)
```
```{r}
# Add a column "ID" with information based on the "Sample" column, basically making a group where reps are within a sample to calculate mean
STRIPs <- STRIPs %>%
  mutate(ID = ifelse(grepl("B1_manure", Sample), "B1_manure",
                     ifelse(grepl("B2_manure", Sample), "B2_manure",
                            ifelse(grepl("P4", Sample), "Prairie",
                                   ifelse(grepl("P3", Sample), "Border", "Crop")))))
```

### Filtering based on Ct    
For this data, we will exclude Ct values of more than 28 based on experience.        
First we select only rows with Ct less than 28 and then group samples by 
```{r}
hits <- STRIPs %>%
  select(Sample, ID, Assay, Ct) %>%
  filter(Ct <= 28) %>%
  group_by(ID, Assay) %>%
  summarize(mean_ct = mean(Ct, na.rm = F)) %>%
  arrange(ID, Assay, mean_ct)
```

# Wafergen results in soil without manure
```{r}
hits_soil <- hits %>%
  filter(ID %in% c("Prairie", "Border", "Crop"))
p <- ggplot(data=hits_soil, aes(x=ID, y=mean_ct, fill=Assay)) +
geom_bar(stat="identity", color="black", position=position_dodge())+
  theme_minimal()
p
```
  
# Wafergen results in manure
```{r}
hits_manure <- hits %>%
  filter(ID %in% c("B1_manure", "B2_manure"))
p <- ggplot(data=hits_manure, aes(x=ID, y=mean_ct, fill=Assay)) +
geom_bar(stat="identity", color="black", position=position_dodge())+
  theme_minimal()
p
```    
   
## tet genes
```{r}
tet <- filter(hits, grepl("tet", Assay)) 
tetp <- ggplot(data=tet, aes(x=ID, y=mean_ct, fill=Assay)) +
geom_bar(stat="identity", color="black", position=position_dodge())+
  theme_minimal()
tetp
```
    
## erm genes
```{r}
erm <- filter(hits, grepl("erm", Assay)) 
ermp <- ggplot(data=erm, aes(x=ID, y=mean_ct, fill=Assay)) +
geom_bar(stat="identity", color="black", position=position_dodge())+
  theme_minimal()
ermp
```

## sul genes
```{r}
sul <- filter(hits, grepl("sul", Assay)) 
sulp <- ggplot(data=sul, aes(x=ID, y=mean_ct, fill=Assay)) +
geom_bar(stat="identity", color="black", position=position_dodge())+
  theme_minimal()
sulp
```
