---
title: "Analysis of WORLE 2017 soils and water 16s rRNA"
author: "Jared"
date: "6/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### First we read in the data, a R data structure of the phyloseq object       
```{r, echo = F}
library(phyloseq)
library(tidyverse)
library(gplots)
library(phylosmith)
```
```{r}
phy <- readRDS("./data/RDS/worle_with_meta.RDS")
colnames(sample_data(phy))
unique(phy@sam_data$day)
```

### Get a list of ASVs from manure and soil before manure was applied
```{r}
manure_otus <- subset_samples(phy, matrix == "manure") %>%
  filter_taxa(function(x) sum(x) >= 1, T) %>%
  otu_table() %>%
  rownames()

#aa <- subset_samples(phy, matrix %in% c("manure", "soil")) %>%
#  unique_taxa(treatment = c("matrix"))
  
cc <- subset_samples(phy, treatment == "no_manure_strip" | day == "b") %>%
  filter_taxa(function(x) sum(x) >= 1, T) %>%
  otu_table() %>%
  rownames()



vvv_diag <- venn(list("Manure_ASVs" = manure_otus, "Soil_ASVs" = cc))
persitors <- attr(vvv_diag, "intersections")$Manure_ASVs
# See if lines 32 to 40 make the same plot as using the chunk below, don't change the chunk below
```

### Find ASVs from manure that are not present in any soil baseline sample
```{r}
v_diag <- venn(list("Manure_ASVs" = aa$manure, "Soil_ASVs" = aa$soil))

just_manure_ASVs <- attr(vvv_diag, "intersections")$Manure_ASVs
```

### Plot relative abundance of only manure associated ASVs   
```{r}
phy_soil <- subset_samples(phy, matrix == "soil") %>%
  filter_taxa(function(x) sum(x) >= 1, T)

my_subset <- subset(otu_table(phy_soil), rownames(otu_table(phy_soil)) %in% persitors)

manure_ASVs_physeq <- merge_phyloseq(my_subset, tax_table(phy), sample_data(phy)) %>%
  filter_taxa(function(x) sum(x) >= 1, T) 
manure_ASVs_physeq 
```

### Trying to get a good plot of the abundance of manure associated ASVs
```{r}
taxa_abundance_bars(manure_ASVs_physeq, classification = "Genus", treatment = c("treatment"), subset = c("manured_control", "no_manure_strip", "manured_strip"), transformation = "mean", wrap_by = "day")
```
## Working off my own scripts from the incubation manuscript   
We would like to generate a plot showing only soil samples from `S1-S5` to determine if proximity to strips affects persistance. Control plots are the greatest distance from where manure was applied. 

```{r}
#Rename treatments to more informative titles
data <- data.frame(sample_data(manure_ASVs_physeq)) %>%
  mutate(TreatmentAndDay = paste(treatment, day))

rownames(data) <- data$id
sample_data(manure_ASVs_physeq) <- data
sample_data(manure_ASVs_physeq)$day <- as.factor(sample_data(manure_ASVs_physeq)$day)
sample_data(manure_ASVs_physeq)
unique(manure_ASVs_physeq@sam_data$day)
```

```{r}
# test conglomorate samples from phylosmith
tt <- conglomerate_samples(manure_ASVs_physeq, c("TreatmentAndDay")) # slow and still gives NA, never completed testing this

manure_ASVs_soil <- subset_samples(manure_ASVs_physeq, matrix %in% c("soil")) %>%
  filter_taxa(function(x) sum(x) >= 1, T) %>%
  melt_phyloseq() %>%
  arrange(Phylum)

test <- manure_ASVs_soil %>%
  group_by(TreatmentAndDay) 

test
```

