---
title: "WORLE rainfall simulation ordinations"
output:
  html_document:
    df_print: paged
    fig_width: 8.5
    fig_height: 11
    fig_caption: true
    toc: true
    toc_float: true
    number_sections: false
    code_folding: hide
---

```{r, echo=F}
library(phyloseq)
library(phylosmith)
library(tidyverse)
library(viridis)
library(vegan)
```
```{r, message = F, warning = F}
phy <- readRDS("./data/RDS/worle_with_meta.RDS")

data <- data.frame(sample_data(phy)) %>%
  filter(matrix == "soil") %>%
  mutate(Vegetation = case_when(in_plot_location %in% c("s1", "s2", "s3", "s4") ~ "Crop",
                                in_plot_location %in% c("s6", "s7", "s8", "s9") ~ "Prairie",
                                TRUE ~ "Border"))

all_data <- data.frame(sample_data(phy)) %>%
  full_join(data)

all_data <- all_data %>%
  mutate(day = replace(day, day == 't2', "t02"))

rownames(all_data) <- all_data$id

sample_data(phy) <- all_data
```

# Measuring differences in bacterial composition
## Depth 1
Visualizing with ordinations is one method of describing bacterial community variations, similarity, and dissimilarity.   

The first question to ask is if crop soils are different from prairie soils, we will look at depth one of "no_manure_strip" treatment. This question is obviously more focused on elucidating the differences in crop and prairie in agroecosystems with PFS recently planted. This does not consider manure treatment. 

```{r}
# Subset to desired samples and remove ASVs observed once
PvCd1 <- subset_samples(phy, matrix == "soil" & treatment == "no_manure_strip" & depth == "d1" & !in_plot_location == "s5") %>%
  filter_taxa(function(x) sum(x) >= 1, T)
```

### PCoA with bray
Calculate dissimilarity between soil vegetation samples from depth 1 of no_manure_strip and plot the ordination using "bray-curtis"  

```{r}
phy.ord.bray <- ordinate(PvCd1, method = "PCoA", distance = "bray")
plot_ordination(PvCd1, phy.ord.bray, color = "Vegetation") + 
  stat_ellipse(geom = "polygon", type = "norm", alpha = 0.1, aes(fill = Vegetation)) +
  scale_colour_viridis_d() + ggplot2::theme_bw() +
  guides(fill=guide_legend(title="Vegetation"))

ord_data <- data.frame(sample_data(PvCd1))

dist <- phyloseq::distance(PvCd1, method = "bray")
adonis2(dist ~ Vegetation, data = ord_data)
```

### tsne with bray
```{r}
tsne_phyloseq(PvCd1, treatment = c('Vegetation'), perplexity = 10, circle = TRUE, colors = 'default') +
  scale_fill_viridis(discrete = T, option = "viridis") + ggplot2::theme_bw() +
  guides(fill=guide_legend(title="Vegetation"))
```

### Different distance matrices  
Hellinger transformation using vegan, this transformation gives low weights to variables with low counts and many zeros. The transformation itself comprises dividing each value in a data matrix by its row sum, and taking the square root of the quotient. More information on this can be found here:  <a href="https://mb3is.megx.net/gustame/reference/transformations">GUSTA ME</a>   

```{r}
# source functions
# contains function for converting to vegan OTU table
source("code/functions.R")
```

```{r}
dis <- vegdist(decostand(veganotu(PvCd1), "hell"), "euclidean")
ord.dis <- ordinate(PvCd1, method = "RDA", distance = dis)
plot_ordination(PvCd1, ord.dis, color = "Vegetation") + 
  stat_ellipse(geom = "polygon", type = "norm", alpha = 0.1, aes(fill = Vegetation)) +
  scale_colour_viridis_d() + ggplot2::theme_bw() +
  guides(fill=guide_legend(title="Vegetation"))
adonis2(dis ~ Vegetation, data = data.frame(sample_data(PvCd1)))
```

## Depth 2
```{r}
# Subset to desired samples and remove ASVs observed once
PvCd2 <- subset_samples(phy, matrix == "soil" & treatment == "no_manure_strip" & depth == "d2" & !in_plot_location == "s5") %>%
  filter_taxa(function(x) sum(x) >= 1, T)
```

### PCoA with bray
```{r}
phy.ord.bray2 <- ordinate(PvCd2, method = "PCoA", distance = "bray")
plot_ordination(PvCd2, phy.ord.bray2, color = "Vegetation") + 
  stat_ellipse(geom = "polygon", type = "norm", alpha = 0.1, aes(fill = Vegetation)) +
  scale_colour_viridis_d() + ggplot2::theme_bw() +
  guides(fill=guide_legend(title="Vegetation"))
adonis2(phyloseq::distance(PvCd2, method = "bray") ~ Vegetation, data = data.frame(sample_data(PvCd2)))
```

### tsne with bray
```{r}
tsne_phyloseq(PvCd2, treatment = c('Vegetation'), perplexity = 10, circle = TRUE, colors = 'default') +
  scale_fill_viridis(discrete = T, option = "viridis") + ggplot2::theme_bw() +
  guides(fill=guide_legend(title="Vegetation"))
```

### Hellinger transformed PCA on depth 2
```{r}
dis2 <- vegdist(decostand(veganotu(PvCd2), "hell"), "euclidean")
ord.dis2 <- ordinate(PvCd2, method = "RDA", distance = dis2)
plot_ordination(PvCd2, ord.dis2, color = "Vegetation") + 
  stat_ellipse(geom = "polygon", type = "norm", alpha = 0.1, aes(fill = Vegetation)) +
  scale_colour_viridis_d() + ggplot2::theme_bw() +
  guides(fill=guide_legend(title="Vegetation"))
adonis2(dis2 ~ Vegetation, data = data.frame(sample_data(PvCd2)))
```