---
title: "WORLE rainfall simulation ordinations"
output:
  html_document:
    df_print: paged
    fig_width: 8.5
    fig_height: 11
    fig_caption: true
    toc: true
    toc_float: true
    number_sections: false
    code_folding: hide
---

```{r, echo=F}
library(phyloseq)
library(phylosmith)
library(tidyverse)
library(viridis)
```
```{r, message = F, warning = F}
phy <- readRDS("../../data/RDS/worle_with_meta.RDS")
```

# TSNE plots   

## The whole experiment TSNE   
```{r, message = F, warning = F}
by_matrix_tsne <- tsne_phyloseq_ggplot(phy, treatment = c('matrix'), perplexity = 10, circle = TRUE, colors = 'default') +
  scale_fill_viridis(discrete = T, option = "viridis") + ggplot2::theme_bw() +
  guides(fill=guide_legend(title="Matrix"))

by_matrix_tsne
```

We can see that each matrix clusters seperately from each other, with the water and manure being very similar. This is to be expected as the water is runoff from a rainfall simulation performed immediately after the manure was applied.    

#### Removing ASVs observed less than 15 times, DADA2 introduces many singletons which adds "noise" to ordinations making clustering less obvisous than OTU based ordinations.    

## Just water samples mintax 15 
```{r, message = F, warning = F}
water <- subset_samples(phy, matrix == "water") %>%
  filter_taxa(function(x) sum(x) > 15, T)

water_tsne <- tsne_phyloseq_ggplot(water, treatment = c('treatment'), perplexity = 10, circle = T, colors = 'default') +
  scale_fill_viridis(discrete = T, option = "viridis") + ggplot2::theme_bw() +
  guides(fill = guide_legend(title = "Treatment")) 
water_tsne
```

## Just soil samples, depth 1         
```{r, message = F, warning = F}
soil_depth_1 <- subset_samples(phy, matrix == "soil" & depth == "d1") %>%
  filter_taxa(function(x) sum(x) > 15, T) 
tsne_soil_depth_1 <- tsne_phyloseq_ggplot(soil_depth_1, treatment = c('treatment'), perplexity = 10, circle = TRUE, colors = 'default') +
  scale_fill_viridis(discrete = T, option = "viridis") + ggplot2::theme_bw() +
  guides(fill=guide_legend(title="Treatment")) 

tsne_soil_depth_1
```

## Just prairie soil samples, depth 1       
```{r, message = F, warning = F}
only_prairie_samples <- subset_samples(soil_depth_1, in_plot_location %in% c("s6", "s7", "s8", "s9")) %>%
  filter_taxa(function(x) sum(x) > 15, T)
              
only_prairie_samples_tsne <- tsne_phyloseq_ggplot(only_prairie_samples, treatment = c('treatment'), perplexity = 10, circle = TRUE, colors = 'default') +
  scale_fill_viridis(discrete = T, option = "viridis") + ggplot2::theme_bw() +
  guides(fill=guide_legend(title="Treatment")) 
only_prairie_samples_tsne
```

## Just crop soil samples, depth 1   
```{r, message = F, warning = F}

only_crop_samples <- subset_samples(soil_depth_1, in_plot_location %in% c("s1", "s2", "s3", "s4", "s5")) %>%
  filter_taxa(function(x) sum(x) > 15, T)
              

only_crop_samples_tsne <- tsne_phyloseq_ggplot(only_crop_samples, treatment = c('treatment'), perplexity = 10, circle = TRUE, colors = 'default') +
  scale_fill_viridis(discrete = T, option = "viridis") + ggplot2::theme_bw() +
  guides(fill=guide_legend(title="Treatment")) 

only_crop_samples_tsne
```

# Depth 2     
## Just soil samples, depth 2          
```{r, message = F, warning = F}
soil_depth_2 <- subset_samples(phy, matrix == "soil" & depth == "d2") %>%
  filter_taxa(function(x) sum(x) > 15, T)

tsne_soil_depth_2 <- tsne_phyloseq_ggplot(soil_depth_2, treatment = c('treatment'), perplexity = 10, circle = TRUE, colors = 'default') +
  scale_fill_viridis(discrete = T, option = "viridis") + ggplot2::theme_bw() +
  guides(fill=guide_legend(title="Treatment"))

tsne_soil_depth_2
```

## Just prairie soil samples, depth 2      
```{r, message = F, warning = F}
only_prairie_samples <- subset_samples(soil_depth_2, in_plot_location %in% c("s6", "s7", "s8", "s9")) %>%
  filter_taxa(function(x) sum(x) > 15, T)
              
only_prairie_samples_tsne <- tsne_phyloseq_ggplot(only_prairie_samples, treatment = c('treatment'), perplexity = 10, circle = TRUE, colors = 'default') +
  scale_fill_viridis(discrete = T, option = "viridis") + ggplot2::theme_bw() +
  guides(fill=guide_legend(title="Treatment"))

only_prairie_samples_tsne
```

#### Facet by day of prairie soil sample, t2 out of order
```{r, message = F, warning = F}
only_prairie_samples_tsne.f <- tsne_phyloseq_ggplot(only_prairie_samples, treatment = c('treatment'), perplexity = 10, circle = TRUE, colors = 'default') +
  scale_fill_viridis(discrete = T, option = "viridis") + ggplot2::theme_bw() +
  guides(fill=guide_legend(title="Treatment")) +
  facet_wrap(~day)

only_prairie_samples_tsne.f
```

## Just crop soil samples, depth 2      
```{r}
only_crop_samples <- subset_samples(soil_depth_2, in_plot_location %in% c("s1", "s2", "s3", "s4", "s5")) %>%
  filter_taxa(function(x) sum(x) > 15, T)
              
only_crop_samples_tsne <- tsne_phyloseq_ggplot(only_crop_samples, treatment = c('treatment'), perplexity = 10, circle = TRUE, colors = 'default') +
  scale_fill_viridis(discrete = T, option = "viridis") + ggplot2::theme_bw() +
  guides(fill=guide_legend(title="Treatment"))

only_crop_samples_tsne
```

## PCoA unweighted uni-frac
```{r}
phy.ord.bray <- ordinate(phy, method = "PCoA", distance = "bray")
plot_ordination(phy, phy.ord.bray, color = "treatment", shape = "matrix")
```

